Fit by design: Developing substrate-specific seed mixtures for functional dike grasslands

### 2023-03-02 manuscript.Rmd
Update manuscript.Rmd

To measure the success of restoration, we chose three indices. First, we observed the persistence of the sown species, which is a key objective of the restoration measure sowing [@Wilsey.2021]. For this purpose, the presence of the sown species was monitored for three consecutive years. Second, we calculated the Favourable Conservation Status (FCS) which distinguishes characteristic and derived diversity [@Helm.2015]. Characteristic diversity is created by established species which belong to the habitat-specific species pool. In other words, these species are associated with the habitat type of interest, whereas derived species not. Third, the recovery was determined by measuring the successional distance to the reference grasslands [@Rydgren.2019]. To complement these indices, the success of restoration or the difference from desired conditions was evaluated [cf. @Brudvig.2017b]. This was reached by comparing the species composition of the experimental plots with reference sites. We tested following hypotheses after four years of observation:

@@

Under the substrate, a 5 cm thick drainage layer of gravel (0-16 mm) was installed. Soil samples of the three substrates were tested on both expositions by mixing several sub-samples from different plots. The sand admixture changed the soil texture, increased the C/N ratio, reduced calcium carbonate, but did hardly change the pH which were within the weak alkaline range (Table \@ref(tab:substrate)). The pH corresponded to a well-established dike vegetation on the Rhine River [@husicka2003], phosphate and potassium were rather scarce for agricultural soils, but magnesium showed high concentrations [@lfl2022]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and the Londo scale was used [@londo1976]. The establishment rates of species are recorded in Appendix A4. In total, 48 species of the species pool of hay meadows (87%) and 46 (79%) of dry grasslands established by 2021, which are rather good ratios [@hedberg2010]. The species established in 31 ± 22% (mean ± SD) of their sown plots. In total, 274 vascular plant species were detected (Appendix A5). Persistence was derived from the 'species losses' component of the temporal beta-diversity index (TBI; 1 − *B*~sor~) which was calculated by comparing the seed mixtures with the respective species composition of each year using Sørensen dissimilarity [@legendre2019]. The successional distance quantifies the ordination distance of a plot to the average reference site [*d~jt,0~*, @Rydgren.2019].

>>>>

To measure the success of restoration, we chose three indices. First, we observed the persistence of the sown species, which is a key objective of the restoration measure sowing [@Wilsey.2021]. For this purpose, the presence of the sown species was monitored for three consecutive years. Second, we calculated the Favourable Conservation Status (FCS) which distinguishes habitat-characteristic and non-typical derived diversity [@Helm.2015]. Third, the recovery completeness was determined by measuring the successional distance to the reference grasslands [@Rydgren.2019]. To complement these indices, the success of restoration or the difference from desired conditions was evaluated [cf. @Brudvig.2017b]. This was reached by comparing the species composition of the experimental plots with reference sites. We tested following hypotheses after four years of observation:

@@

Under the substrate, a 5 cm thick drainage layer of gravel (0-16 mm) was installed. Soil samples of the three substrates were tested on both expositions by mixing several sub-samples from different plots. The sand admixture changed the soil texture, increased the C/N ratio, reduced calcium carbonate, but did hardly change the pH which were within the weak alkaline range (Table \@ref(tab:substrate)). The pH corresponded to a well-established dike vegetation on the Rhine River [@husicka2003], phosphate and potassium were rather scarce for agricultural soils, but magnesium showed high concentrations [@lfl2022]. The vegetation was surveyed in June or July 2018--2021 [@BraunBlanquet.1964] and the Londo scale was used [@londo1976]. The establishment rates of species are recorded in Appendix A4. In total, 48 species of the species pool of hay meadows (87%) and 46 (79%) of dry grasslands established by 2021, which are rather good ratios [@hedberg2010]. The species established in 31 ± 22% (mean ± SD) of their sown plots. In total, 274 vascular plant species were detected (Appendix A5). The recovery completeness was described by the successional distance which quantifies the distance of a plot to the average reference site in the ordination [*d~jt,0~*, @Rydgren.2019]. Persistence was derived from the 'species losses' component of the temporal beta-diversity index (TBI; 1 − *B*~sor~) which was calculated by comparing the seed mixtures with the respective species composition of each year using Sørensen dissimilarity [@legendre2019]. The Favourable Conservation Status (FCS) is the ratio of characteristic and derived diversity, in our case species richness. Characteristic diversity consists of species which belong to a habitat-specific species pool and derived diversity consists of all other species. The habitat-specific species pool consisted of all sown species and other typical species of mesic and dry grasslands [Appendix A5, @Helm.2015].
